---
title: "Homework_Fri Nov 14"
author: "Reem Naji"
date: "11/13/2025"
embed-resources: true
fig-align: center
format: html
---

```{r}
#| message: false
#| label: Setup 

library(tidyverse)
attrition <- read.csv("attrition_small.csv")

```



```{r}
#| output: false
glimpse(attrition)
```

# Problem 1 
$H_0:$ The company's claim on  `worklifebalance` rating is correct & representative  
$H_a:$ The company's claim on `worklifebalance` rating is **not** correct & representative  
```{r}

obs <- table(attrition$WorkLifeBalance) 
print(obs)
```

```{r}
exp_props <- c("best" = 0.10, "better" = 0.65, "good" = 0.20, "bad" = 0.05)
exp <- exp_props * 250
print(exp)
```

```{r}
chisq.test(obs, p = exp_props)
```

```{r}
# extra setp to confirm!
chisq_stat <- sum((obs - exp)^2 / exp)
print(chisq_stat)
```

Given the p-value of $2.2\times^{-16}$ we have a strong evidence against the  null hypothesis. The company's claimed ratings  are very different from the actual ratings. So this claim is unreasonable in the light of the given sample. 


# Problem 2
$H_0:$ The company's claim on `EducationField` distrubtion is correct & representative  
$H_a:$ The company's claim on `EducationField` distrubtion is **not** correct & representative
```{r}
obs_2 <- table(attrition$EducationField) 
edu_props <- c("life science" = 0.55, "marketing" = 0.20, "technical" = 0.20, "other" = 0.05)
exp_2 <- edu_props * 250
print(exp_2)
```

```{r}
chisq.test(obs_2, p = edu_props)
```
Given the extremely small p-value of $8.707\times^{-12}$ we reject the null. The company's claim on education field distrubtion is not supported or reasonable, and the the difference between the actual and claimed is too large to be explained by random chance. 
